---
title: "Statistical Inference - Course Project - Question 1"
date: '2014-09-21'
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r, echo=FALSE, message=FALSE, cache=FALSE}
require(ggplot2)
require(knitr)

set.seed(987654)
opts_chunk$set(echo=TRUE, message=FALSE, cache=FALSE)
```

### Synopsis
In this simulation, I will investigate the distribution of averages of 40 exponential(0.2)s. I will illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s. Initilize our simulation dataset with $\lambda$=0.2, 40 samples and 1000 rows per sample.
```{r}
lambda <- 0.2; s <- 40; n <- 1000;
dist <- matrix(rexp(s*n, rate=lambda), ncol=s, nrow=n)
distMean <- rowMeans(dist)
```

### 1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
Calculate the expected theoretical center and the center (mean) of `distMean`. 
```{r echo=TRUE}
centerTheoretical <- 1/lambda; centerCalculated <- mean(distMean);
```

```{r  DistCenterHist, fig.width=6, fig.height=3, echo=FALSE}
pl <- ggplot(data=data.frame(distMean), aes(x=distMean)) +
  geom_histogram(binwidth=0.25, colour='white') + 
  #ggtitle('Distribution Mean Frequency') + 
  xlab('Distribution Mean') + 
  scale_x_continuous(breaks = round(seq(min(distMean), max(distMean), by = 0.5),1)) + 
  geom_vline(xintercept=centerCalculated, colour='green', size=1) + 
  ylab('Frequency') +
  theme(title=element_text(size=9),
        axis.text=element_text(size=9),
        axis.title=element_text(size=10))
print(pl)
```

The theoretical center of the distribution is: $1/\lambda$ = 1/0.2 = `r centerTheoretical`. Our calculated center of the distribution, indicated by the green line, is `r centerCalculated`. The two are very close.

### 2. Show how variable it is and compare it to the theoretical variance of the distribution.
```{r Q2}
sdTheoretical <- (1/lambda)*(1/sqrt(s)); sdCalculated <- sd(distMean);
varTheoretical <- sdTheoretical^2; varCalculated <- sdCalculated^2;
```

The theoretical standard deviation is calculated as follow: $(\frac{1}{\lambda} * \frac{1}{\sqrt{n}})$ = `r sdTheoretical` yielding a variance of `r varTheoretical`. Our calculated standard deviation of the distribution is `r sdCalculated` yielding a calculated variance of `r varCalculated`. Again the theoretical and calculated values are very close to each other.

### 3. Show that the distribution is approximately normal.
In an effort to compare our calculated distribution with a normal distribution I render the histogram shown earlier with the Y Axis showing the density as opposed to the previous shown frequency and I then overlay a Normal Distribution (red line) using the our `centerCalculated` and `sdCalculated` variables. This plot clearly shows that our distribution is approximately normal.

```{r  NormHist, fig.width=6, fig.height=2.5, echo=FALSE}
pl <- ggplot(data=data.frame(distMean), aes(x=distMean)) +
  geom_histogram(aes(y = ..density..), binwidth=0.25, colour='white') + 
    #ggtitle('Calculated Distribution vs Normal Distribution') + 
    xlab('Distribution Mean') + 
    scale_x_continuous(breaks = round(seq(min(distMean), max(distMean), by = 0.5),1)) + 
    ylab('Density') +
    stat_function(
      fun = dnorm, 
      args = with(data.frame(distMean), c(mean = mean(distMean), sd = sd(distMean))), 
      colour='red', size=1) +
    theme(title=element_text(size=9),
        axis.text=element_text(size=9),
        axis.title=element_text(size=10))
print(pl)
```

### 4. Evaluate the coverage of the confidence interval for $1/\lambda = \bar{X} \pm 1.96 \frac{S}{\sqrt{n}}$.
Start by creating a function `calcCover` that returns the coverage for a specified lambda, then create a range of lambdas, calculate their respective coverage and lastly calculate the vertical bounds (`vBounds`). Finally plot the `cover` against the `lambdas`.
```{r}
calcCover <- function(x) {
  offset <-  (1.96 * sqrt(1/lambda**2/s));
  ll <- distMean - offset; ul <- distMean + offset;
  mean(ll < x & ul > x)
}
lambdas <- seq(4.25, 5.75, by=0.01); cover <- sapply(lambdas, calcCover);
vBounds <- centerCalculated + (c(-1,1) * 1.96 * sdCalculated/sqrt(s))
```

```{r  Cover, fig.width=6, fig.height=2.5, echo=FALSE, warning=FALSE}
pl <- ggplot(data.frame(lambdas, cover), aes(x = lambdas, y = cover)) + 
  geom_line(size = 1) + 
  geom_vline(xintercept=vBounds, col="red") +
  geom_hline(yintercept = 0.95, col="blue") + 
  xlab('Lambda') +
  ylab('Coverage') +
    ylim(.85, 1.0) +
    theme(title=element_text(size=9),
        axis.text=element_text(size=9),
        axis.title=element_text(size=10))

print(pl)
```

The horizontal blue line marks the 95% confidence interval and the vertical red lines (`r vBounds[1]` and `r vBounds[2]`) marks the above 95% interval. 